---
title: "Data 605 HW 11"
author: "Matthew Lucich"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 

```{r}
summary(cars)
glimpse(cars)
```

#### Create scatterplot and review if relationship is approximately linear

```{r}
plot(cars$speed, cars$dist, xlab="Speed", ylab="Stopping Distance", main="Stopping Distance vs Speed")
```

Relationship looks approximately linear, so we will proceed with fitting a linear regression model.

#### Fitting linear regression model

```{r}
model <- lm(dist ~ speed, data=cars)
summary(model)
```

The summary shows that both intercept and speed are significant at a 95% confidence level. The adjusted R-squared is 0.64 which is moderately high. The F-statistic indicates the model is significant. 


```{r}
plot(model)
```

Residual vs Fitted plot shows there is no serious heteroskedasticity and since there does not appear to be a trend we can likely conclude the residuals are independent. The Q-Q plot shows while there are potentially some influential outliers, the residuals are approximately normal.

Based on our analysis we can be confident the fitted model meets the assumptions of linear regression.




