---
title: "Data 605 HW 15"
author: "Matthew Lucich"
date: "4/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Deriv)
library(Ryacas)
```

### 1) Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

```{r}

independent <- c(5.6, 6.3, 7, 7.7, 8.4)
dependent <- c(8.8, 12.4, 14.8, 18.2, 20.8)

model <- lm(dependent ~ independent)
summary(model)

```

**Solution**:

$y=-14.80 + 4.26x$


### 2) Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form (x, y, z). Separate multiple points with a comma.

$f(x,y) = 24x - 6xy^2 - 8y^3$

```{r}

f=expression(24*x - 6*x*y^2 - 8*y^3)

fx <- D(f,'x')
Simplify(fx)

fy <- D(f,'y')
Simplify(fy)

```

Solve for y

$0 = 24 - 6y^2$

$6y^2 = 24$

$6y^2 = 24$

$y^2 = 4$

$y \pm 2$



Solve for x

$0 = -12y (x+2y)$

$0 = -12yx -24y^2$

$12yx  = -24y^2$

$x = \frac{-24y^2}{12y}$

$x = -2y$

$x = -2(\pm2)$

$x \pm 4$


```{r}


fxx <- D(D(f,'x'),'x')
Simplify(fxx)

fxy <- D(D(f,'x'),'y')
Simplify(fxy)

fyy <- D(D(f,'y'),'y')
Simplify(fyy)

```

$0 * -12(x+4y) - (-12y)^2$

$-144y^2$

Evaluated at with 2 and -2 (x is ignored due to no term) the result is: -576.

Since, -576 is less than zero, (-4, 2) and (4, -2) are saddle points.

There is no local maxima or local minima.


### 3) A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81  21x + 17y units of the "house" brand and 40 + 11x  23y units of the "name" brand.

Step 1. Find the revenue function R ( x, y ).

```{r}

house <- expression(81-21*x+17*y)
name <- expression(40 + 11*x - 23*y)

house_and_name <- expression(x*(81-21*x+17*y) + y*(40 + 11*x - 23*y))

Simplify(house_and_name)

```

Expanding out...
$17xy + 81x - 21x^2 + 11xy + 40y - 23y^2$

Revenue function is:
$28xy + 81x - 21x^2 + 40y - 23y^2$


Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}

x <- 2.3
y <- 4.1

x*(81 - 21*x + 17*y) + y*(40 + 11*x - 23*y)

```



### 4) A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by C(x, y) = 1/6 x2 + 1/6 y2 + 7x + 25y + 700, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

Cost function is: 

$\frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$

Required output per week can be defined as:
$x + y = 96$

Which we can rewrite as:
$x = 96 - y$

Let's substitute the expression above for x:

$\frac{1}{6} (96 - y)^2 + \frac{1}{6} y^2 + 7(96 - y) + 25y + 700$

Which simplifies to:

$1536 - 32y + \frac{1}{6}y^2 + \frac{1}{6} y^2 + 672 - 7y + 25y + 700$

$2908 - \frac{1}{3}y^2 - 14y$


Let's take the first derivative to find minimum cost...

```{r}

f <- expression(2908 - (1/3)*y^2-14*y)
fx <- D(f,'y')
Simplify(fx)

```

$0 = -\frac{2}{3}y - 14$

Solve for y:

$y = 21$

Therefore, if 21 units should be produced in Denver..

$x + 21 = 96$

Then, the amount of units produced in Los Angeles should be:

$x = 75$



### 5) Evaluate the double integral on the given region. Write your answer in exact form without decimals.

$\int_{2}^{4} \int_{2}^{4} e^{8x+3y} \,dx dy$

Apply u-substitution

u = $8x+3y$

Take derivative with respect to x:

$8+0$

$\frac{du}{dx} = 8+0$

$dx = \frac{1}{8}du$

Now plug in 2 and 4 to u for adjusted boundaries:

$8(2) + 3y = 16 + 3y$

$8(4) + 3y = 32 + 3y$

The updated double integral becomes:

$\int_{16 + 3y}^{32 + 3y} e^u \frac{1}{8} \,du$

Move the constant to the front:

$\frac{1}{8} \int_{16 + 3y}^{32 + 3y} e^u \,du$


$\frac{1}{8} (e^{32 + 3y} - e^{16 + 3y}) \,du$


$\int_{2}^{4} \frac{1}{8} (e^{32 + 3y} - e^{16 + 3y}) \,dy$

Apply sum rule

$\frac{1}{8} \int_{2}^{4} e^{32 + 3y} - \int_{2}^{4}e^{16 + 3y} \,dy$


Apply u-substitution

u = $32 + 3y$

Take derivative with respect to y:

$0+3$

$\frac{du}{dy} = 0+3$

$dy = \frac{1}{3}du$

Now plug in 2 and 4 to u for adjusted boundaries:

$32 + 3(2) = 38$

$32 + 3(4) = 44$

The updated double integral becomes:

$\int_{38}^{44} e^u \frac{1}{3} \,du$


$\frac{1}{3} (e^{44} - e^{38})$


Apply u-substitution

u = $16 + 3y$

Take derivative with respect to y:

$0+3$

$\frac{du}{dy} = 0+3$

$dy = \frac{1}{3}du$

Now plug in 2 and 4 to u for adjusted boundaries:

$16 + 3(2) = 22$

$16 + 3(4) = 28$

The updated double integral becomes:

$\int_{22}^{28} e^u \frac{1}{3} \,du$


$\frac{1}{3} (e^{28} - e^{22})$


Now combine:

$(e^{44} - e^{38}) - (e^{28} - e^{22})$

$e^{44} - e^{38} - e^{28} + e^{22}$


Insert the constants:

$\frac{1}{8}(\frac{1}{3}(e^{44} - e^{38} - e^{28} + e^{22}))$


