---
title: "Data 605 HW 13"
author: "Matthew Lucich"
date: "4/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1) Use integration by substitution to solve the integral below.

$\int 4 e^{-7x}dx$

Move constant

$4 \int e^{-7x}dx$

Use u-substitution

$u = -7x$

Solve for dx

$\frac{du}{dx} = -7$

$dx = -\frac{1}{7}du$

Replace dx

$4 \int -\frac{1}{7} e^{u}du$

Substitute u back in and add constant

$-\frac{4}{7}e^{-7x}+C$


### 2) Biologists are treating a pond contaminated with bacteria...

$\frac{dN}{dt} = \frac{3150}{t^4} - 220$

$\int \frac{3150}{t^4} - 220 dt$

$N' = 3150 \frac{t^{-3}}{-3} -220t + c$

Set equal to 6530 and substitute 1

$6530 = -1050  (1^{-3}) -220(1) + c$

$6530 = -1270 + c$

$7800 = c$

Therefore, the function is:

$\frac{-1050}{t^3} -220t + 7800$



### 3) Find the total area of the red rectangles in the figure below, where the equation of the line is f ( x ) = 2x  9.


```{r}

func <- function(x) {2*x-9}
integrate(func, lower = 4.5, upper = 8.5)

```



### 4) Find the area of the region bounded by the graphs of the given equations.

Set equations equal to each other and solve for x:

$x^2-2x-2=x+2$

$x^2-3x-4$

$(x+1)(x-4)$

Therefore, we integrate -1 to 4, the difference of the equations

```{r}

func <- function(x) {x+2 - (x^2-2*x-2)}
integrate(func, lower = -1, upper = 4)

```


### 5) A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

**Holding cost**

$H(x) = (holding-cost-per-unit) * \frac{x}{2}$

$3.75  * (x/2)$

$1.875x$

**Reorder cost**

$R(x) = (cost-per-order) * \frac{total-units}{x}$

$8.25 * (110/x)$

$\frac{907.5}{x}$

**Total inventory cost**

$C(x) = H(x) + R(x)$

$1.875x + \frac{907.5}{x}$


Find first derivative of total inventory cost function:

```{r}

f=expression(1.875*x + (907.5/x))
D(f,'x')

```

Set equal to zero and solve for x:

$0 = 1.875 + \frac{907.5}{x^2}$

$x^2 = 484$

$x\pm22$

Therefore, since -22 is not in our range, the lot size that will minimize costs is:

**Lot size: 22**

Which means the number of orders that minimizes costs is:

**Orders: 5** (110/22)



### References

[Minimizing Inventory Costs](https://www.youtube.com/watch?v=EbNQZqsQyz0)
