---
title: "Data 605 HW 14"
author: "Matthew Lucich"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Deriv)
```

## This week, we’ll work out some Taylor Series expansions of popular functions.


### $f(x) = \frac{1}{(1-x)}$

```{r}

f=expression(1/(1-x))

# First derivative
derv_1 <- D(f,'x')
Simplify(derv_1)
derv_1_func <- function(x) {1/(1 - x)^2}
# Evaluate at zero
derv_1_func(0)

# Second derivative
derv_2 <- D(D(f,'x'),'x')
Simplify(derv_2)
derv_2_func <- function(x) {2/(1 - x)^3}
# Evaluate at zero
derv_2_func(0)

# Third derivative
derv_3 <- D(D(D(f,'x'),'x'),'x')
Simplify(derv_3)
derv_3_func <- function(x) {6/(1 - x)^4}
# Evaluate at zero
derv_3_func(0)

# Fourth derivative
derv_4 <- D(D(D(D(f,'x'),'x'),'x'),'x')
Simplify(derv_4)
derv_4_func <- function(x) {24/(1 - x)^5}
# Evaluate at zero
derv_4_func(0)

# Fifth derivative
derv_5 <- D(D(D(D(D(f,'x'),'x'),'x'),'x'),'x')
Simplify(derv_5)
derv_5_func <- function(x) {120/(1 - x)^6}
# Evaluate at zero
derv_5_func(0)

```

We observe that the denominator, n!, yields the same result as the numerator:

```{r}

factorial(1)
factorial(2)
factorial(3)
factorial(4)
factorial(5)

```

Therefore, the Taylor Series simplifies to:

$X^n$



### $f(x) = e^x$

```{r}

f=expression(e^x)

# First derivative
derv_1 <- D(f,'x')
Simplify(derv_1)
derv_1_func <- function(x) {exp(x)* log(exp(1))}
# Evaluate at zero
derv_1_func(0)

# Second derivative
derv_2 <- D(D(f,'x'),'x')
Simplify(derv_2)
derv_2_func <- function(x) {exp(x) * log(exp(1))^2}
# Evaluate at zero
derv_2_func(0)

# Third derivative
derv_3 <- D(D(D(f,'x'),'x'),'x')
Simplify(derv_3)
derv_3_func <- function(x) {exp(x) * log(exp(1))^3}
# Evaluate at zero
derv_3_func(0)

# Fourth derivative
derv_4 <- D(D(D(D(f,'x'),'x'),'x'),'x')
Simplify(derv_4)
derv_4_func <- function(x) {exp(x) * log(exp(1))^4}
# Evaluate at zero
derv_4_func(0)

# Fifth derivative
derv_5 <- D(D(D(D(D(f,'x'),'x'),'x'),'x'),'x')
Simplify(derv_5)
derv_5_func <- function(x) {exp(x) * log(exp(1))^5}
# Evaluate at zero
derv_5_func(0)

```

Therefore, we see a polynomial approximation is:

```{r}

sum((1/factorial(1)) + (1/factorial(2)) + (1/factorial(3)) + (1/factorial(4)) + (1/factorial(5)))

```

More formally, the Taylor Series can be represented as:

$\sum_{n=0}^{\infty} \frac{x^n}{n!}$



### $f(x) = ln(1+x)$


```{r}

f=expression(log(1+x))

# First derivative
derv_1 <- D(f,'x')
Simplify(derv_1)
derv_1_func <- function(x) {1/(1 + x)}
# Evaluate at zero
derv_1_func(0)

# Second derivative
derv_2 <- D(D(f,'x'),'x')
Simplify(derv_2)
derv_2_func <- function(x) {-(1/(1 + x)^2)}
# Evaluate at zero
derv_2_func(0)

# Third derivative
derv_3 <- D(D(D(f,'x'),'x'),'x')
Simplify(derv_3)
derv_3_func <- function(x) {2/(1 + x)^3}
# Evaluate at zero
derv_3_func(0)

# Fourth derivative
derv_4 <- D(D(D(D(f,'x'),'x'),'x'),'x')
Simplify(derv_4)
derv_4_func <- function(x) {-(6/(1 + x)^4)}
# Evaluate at zero
derv_4_func(0)

# Fifth derivative
derv_5 <- D(D(D(D(D(f,'x'),'x'),'x'),'x'),'x')
Simplify(derv_5)
derv_5_func <- function(x) {24/(1 + x)^5}
# Evaluate at zero
derv_5_func(0)

```


Therefore, we see a polynomial approximation is:

```{r}

sum((1/factorial(1)) + (-1/factorial(2)) + (2/factorial(3)) + (-6/factorial(4)) + (24/factorial(5)))

```

The approximation can also be written as:

$0 + \frac{1}{1!}x - \frac{1}{2!}x^2 + \frac{2}{3!}x^3 - \frac{6}{4!}x^3$


More formally, the Taylor Series can be represented as:

$\sum_{n=1}^{\infty} -1^{n+1} \frac{x^n}{n}$



